Non-kissing and non-crossing complexes \\ for gentle algebras

### 2018-05-30 draft/accordionComplexSurfaces.tex
minor changes

\usepackage[T1]{fontenc}
\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap,  shuffle, xargs, multicol, overpic, pdflscape, multirow, hvfloat, minibox, accents, array, xifthen, ae, aecompl, blkarray, pifont, mathtools, a4wide}
\usepackage{marginnote}
\hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue}
\usepackage[all]{xy}
\usepackage[bottom]{footmisc}
\usepackage{tikz}
\usepackage{tkz-graph}
\usetikzlibrary{trees, decorations, decorations.markings, shapes, arrows, matrix, calc, fit, intersections, patterns, angles}
\graphicspath{{figures/}}
\makeatletter\def\input@path{{figures/}}\makeatother
\usepackage{caption}
\captionsetup{width=\textwidth}

@@

% accordion and zigzag complexes
\newcommand{\surface}{\mathcal{S}} % an orientable surface
\newcommand{\dual}{^*} % duality
\newcommand{\dissection}{\mathrm{D}} % dissection of a marked surface
\newcommand{\faces}{\mathcal{F}} % faces of a dissection
\renewcommandx{\AC}[1][1=\dissection]{\mathcal{K}_{\mathrm{acc}}(#1)} % accordion complex
\newcommandx{\ZC}[1][1=\dissection\dual]{\mathcal{K}_{\mathrm{zz}}(#1)} % zigzag complex
\newcommandx{\NCC}[1][1={\dissection, \dissection\dual}]{\mathcal{K}_{\mathrm{nc}}(#1)} % non-crossing complex

@@

\begin{abstract}
We consider three combinatorial models for the support $\tau$-tilting complex of a gentle bound quiver: the accordion complex, the zigzag complex, and the non-kissing complex. We unify these models in the same picture.
\end{abstract}

@@

Possible titles, depending on what we want to insist on:
\begin{enumerate}
\item Accordion / zigzag / non-kissing / support $\tau$-tilting complexes all coincide for gentle algebras
\item Non-kissing and non-crossing complexes for gentle algebras
\item Combinatorial models for support $\tau$-tilting on gentle algebras: accordion, zigzag and non-kissing complexes
\end{enumerate}

@@

\section{Accordion complex, zigzag complex, and non-crossing complex}

@@

\begin{definition}
A \defn{marked surface}~$\bar\surface \eqdef (\surface,M)$ is an orientable surface~$\surface$ with boundaries, together with a set~$M$ of marked points which can be on the boundary of~$\surface$ or not.
An \defn{arc} on~$\bar\surface$ is a curve on~$\surface$ connecting two marked points of~$M$ and whose interior is disjoint from~$M$ and the boundary of~$\surface$ (note that arcs homotopic to boundary arcs are allowed here).
As usual, arcs are considered up to homotopy relative to their endpoints in~$\surface \ssm M$.
Two arcs \defn{cross} when they intersect in their interior.
\end{definition}

@@

\subsection{Accordion complex, zigzag complex, and non-crossing complex}

@@

\begin{definition}
\label{def:accordion}
A \defn{$\dissection$-accordion} is an arc~$\gamma$ of~$(\surface, B)$ such that any two arcs of $\dissection$ crossed consecutively by~$\gamma$
share a marked point with the property that the region bounded by~$\gamma$ and the two arcs around the marked point is homeomorphic to a disk
(in particular, it contains no punctures).
\yann{Il faut probablement ajouter des conditions aux bords}
\end{definition}

@@

\begin{definition}
\label{def:zigzag}
A \defn{$\dissection\dual$-zigzag} is an arc~$\alpha$ of~$(\surface, B)$ such that, for each pair of faces~$(f,g) \in \faces(\dissection\dual)$ that~$\alpha$ enters consecutively by crossing~$\beta\in\dissection\dual$,~$\alpha$ cuts the union of the two cells $f$ and $g$ glued along~$\beta$ into two components, one containing the marked point $f\dual$ and the other containing $g\dual$.
\end{definition}

\begin{definition}
\label{def:zigzagComplex}
The \defn{$\dissection\dual$-zigzag complex}~$\ZC$ is the simplicial complex of pairwise non-crossing $\dissection$-zigzags.
\vincent{Reduced zigzag complex.}
\end{definition}

\begin{remark}
Note that the $\dissection$-accordions do not depend on~$\dissection\dual$, only on~$\dissection$ (and~$M\dual$ which is determined by~$\dissection$ by Proposition~\ref{prop:conditionsDualDissections}).
Similarly, the $\dissection\dual$-zigzags do not depend on~$\dissection$.
\end{remark}

\begin{proposition}
\label{prop:accordionsZigzags}
For two dual cellular dissections~$\dissection$ and~$\dissection\dual$, the $\dissection$-accordions are precisely the $\dissection\dual$-zigzags and the $\dissection$-accordion complex coincides with the $\dissection\dual$-zigzag complex.
\end{proposition}

@@

Case 2: The arc~$\alpha$ consecutively crosses two arcs~$\beta$ and~$\gamma$ of~$\dissection$ that share a unique common endpoint, but the region delimited by~$\alpha$,~$\beta$ and~$\gamma$ is not a disk.
Since the dissection~$\dissection$ is cellular, that region has to contain some marked point~$m\in\dissection\dual$.
One conclude that~$\alpha$ is not a~$\dissection\dual$-zigzag similarly as in the first case, by considering the component of~$m\dual\setminus\alpha$ that do not contain~$m$.

@@

\begin{definition}
\label{def:noncrossingComplex}
We call \defn{non-crossing complex} of the pair of dual cellular dissections~$(\dissection, \dissection\dual)$ the simplicial complex~$\NCC = \AC = \ZC$.
\end{definition}

\begin{remark}
To clearly visualize Proposition~\ref{prop:accordionsZigzags}, we follow the convention to draw the $\dissection$-accordions (resp.~$\dissection\dual$-zigzags) so that they intersect the arcs of~$\dissection$ (resp.~of~$\dissection\dual$) only at the intersection points of the arcs of~$\dissection$ with the arcs of~$\dissection\dual$.
\vincent{That is interior points}
\end{remark}

\begin{remark}
The $\dissection$-zigzags and the $\dissection\dual$-accordions are defined dually and also coincide.
\end{remark}

\begin{remark}
A quick disclaimer about accordion complexes and zigzag complexes of arbitrary dissections.
As stated in Proposition~\ref{prop:conditionsDualDissections}, a dissection~$\dissection$ with a face containing more than one boundary edge does not admit a dual cellular dissection~$\dissection\dual$.
This prevents to use the definitions given in this section.
Although its accordion complex (resp.~zigzag complex) could still be defined as in~\cite{MannevillePilaud-accordion} (resp.~\cite{GarverMcConville}) for the disk, the resulting complexes would be join of smaller accordion complexes (resp.~zigzag complexes) defined in this paper.
See \cite[Prop.~2.4]{MannevillePilaud-accordion} for a detailed statement in the case of the disk.
\end{remark}

@@

In this section, we show that non-kissing complexes and non-crossing complexes actually coincide.
Namely, any pair of dual cellular dissections~$(\dissection, \dissection\dual)$ defines a gentle bound quiver~$\bar Q_{\dissection, \dissection\dual}$ such that the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.
Conversely, any gentle bound quiver~$\bar Q$ gives rise to a surface~$\surface_{\bar Q}$ equipped with a pair of dual cellular dissections~$\dissection_{\bar Q}$ and~$\dissection\dual_{\bar Q}$ such that the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, \dissection\dual_{\bar Q}]$.
In fact, we show that the accordion, zigzag, and non-kissing complexes can all be pictured in a unique way on the surface.
Interestingly, via the correspondence between pairs of dual cellular dissections and gentle bound quivers, the duality between the dissections~$\dissection$ and~$\dissection\dual$ translates to the Koszul duality between the gentle algebras of~$\bar Q$ and~$\bar Q\dual$.

@@

\subsection{Accordion, zigzag and non-kissing complexe coincide}

@@

\begin{theorem}
\label{thm:complexesCoincide}
\begin{itemize}
\item For any gentle bound quiver~$\bar Q$, the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, \dissection\dual_{\bar Q}]$, therefore to the accordion complex~$\AC[\dissection_{\bar Q}]$ and to the zigzag complex~$\ZC[\dissection\dual_{\bar Q}]$.
\item For any pair of dual cellular dissections~$\dissection, \dissection\dual$ of an oriented surface, the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.
\end{itemize}
\end{theorem}

>>>>

\usepackage[T1]{fontenc}
\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, xargs, multicol, overpic, pdflscape, multirow, hvfloat, minibox, accents, array, xifthen, ae, aecompl, blkarray, pifont, mathtools, a4wide}
\usepackage{marginnote}
\hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue}
\usepackage[all]{xy}
\usepackage[bottom]{footmisc}
\usepackage{tikz}
\usepackage{tkz-graph}
\usetikzlibrary{trees, decorations, decorations.markings, shapes, arrows, matrix, calc, fit, intersections, patterns, angles}
\graphicspath{{figures/}}
\makeatletter\def\input@path{{figures/}}\makeatother
\usepackage{caption}
\captionsetup{width=\textwidth}

@@

% accordion and slalom complexes
\newcommand{\surface}{\mathcal{S}} % an orientable surface
\newcommand{\dual}{^*} % duality
\newcommand{\dissection}{\mathrm{D}} % dissection of a marked surface
\newcommand{\faces}{\mathcal{F}} % faces of a dissection
\renewcommandx{\AC}[1][1=\dissection]{\mathcal{K}_{\mathrm{acc}}(#1)} % accordion complex
\newcommandx{\SC}[1][1=\dissection\dual]{\mathcal{K}_{\mathrm{sla}}(#1)} % slalom complex
\newcommandx{\NCC}[1][1={\dissection, \dissection\dual}]{\mathcal{K}_{\mathrm{nc}}(#1)} % non-crossing complex

@@

\begin{abstract}
We consider three combinatorial models for the support $\tau$-tilting complex of a gentle bound quiver: the accordion complex, the slalom complex, and the non-kissing complex. We unify these models in the same picture.
\end{abstract}

@@

Possible titles, depending on what we want to insist on:
\begin{enumerate}
\item Accordion / slalom / non-kissing / support $\tau$-tilting complexes all coincide for gentle algebras
\item Non-kissing and non-crossing complexes for gentle algebras
\item Combinatorial models for support $\tau$-tilting on gentle algebras: accordion, slalom and non-kissing complexes
\end{enumerate}

@@

\section{Accordion complex, slalom complex, and non-crossing complex}

@@

\begin{definition}
A \defn{marked surface}~$\bar\surface \eqdef (\surface, M)$ is an orientable surface~$\surface$ with boundaries, together with a set~$M$ of marked points which can be on the boundary of~$\surface$ or not.
An \defn{arc} on~$\bar\surface$ is a curve on~$\surface$ connecting two marked points of~$M$ and whose interior is disjoint from~$M$ and the boundary of~$\surface$ (note that arcs homotopic to boundary arcs are allowed here).
As usual, arcs are considered up to homotopy relative to their endpoints in~$\surface \ssm M$.
Two arcs \defn{cross} when they intersect in their interior.
\end{definition}

@@

\subsection{Accordion complex, slalom complex, and non-crossing complex}

@@

\begin{definition}
\label{def:accordion}
A \defn{$\dissection$-accordion} is an arc~$\gamma$ of~$(\surface, B)$ such that whenever~$\gamma$ crosses a face~$f$ of~$\dissection$, it crosses two edges of~$f$ that share a marked point of~$v \in M$ in such a way that~$v$ and~$f\dual$ lie on opposite sides of~$\gamma$.

% the two edges of~$f$ crossed by~$\gamma$ share a marked point~$v \in M$, and~$v$ and~$f\dual$ are on opposite sides of~$\gamma$ in~$f$.
% any two arcs of $\dissection$ crossed consecutively by~$\gamma$ share a marked point with the property that the region bounded by~$\gamma$ and the two arcs around the marked point does not contain.
\end{definition}

@@

\begin{definition}
\label{def:slalom}
A \defn{$\dissection\dual$-slalom} is an arc~$\alpha$ of~$(\surface, B)$ such that, for each pair of faces~$(f,g) \in \faces(\dissection\dual)$ that~$\alpha$ enters consecutively by crossing~$\beta\in\dissection\dual$,~$\alpha$ cuts the union of the two cells $f$ and $g$ glued along~$\beta$ into two components, one containing the marked point $f\dual$ and the other containing $g\dual$.
\end{definition}

\begin{definition}
\label{def:slalomComplex}
The \defn{$\dissection\dual$-slalom complex}~$\SC$ is the simplicial complex of pairwise non-crossing $\dissection$-slaloms.
\vincent{Reduced slalom complex.}
\end{definition}

%\begin{remark}
%Note that the $\dissection$-accordions do not depend on~$\dissection\dual$, only on~$\dissection$ (and~$M\dual$ which is determined by~$\dissection$ by Proposition~\ref{prop:conditionsDualDissections}).
%Similarly, the $\dissection\dual$-slaloms do not depend on~$\dissection$.
%\end{remark}

\begin{proposition}
\label{prop:accordionsSlaloms}
For two dual cellular dissections~$\dissection$ and~$\dissection\dual$, the $\dissection$-accordions are precisely the $\dissection\dual$-slaloms and the $\dissection$-accordion complex coincides with the $\dissection\dual$-slalom complex.
\end{proposition}

@@

Case 2: The arc~$\alpha$ consecutively crosses two arcs~$\beta$ and~$\gamma$ of~$\dissection$ that share a unique common endpoint, but the region delimited by~$\alpha$,~$\beta$ and~$\gamma$ is not a disk.
Since the dissection~$\dissection$ is cellular, that region has to contain some marked point~$m\in\dissection\dual$.
One conclude that~$\alpha$ is not a~$\dissection\dual$-slalom similarly as in the first case, by considering the component of~$m\dual\setminus\alpha$ that do not contain~$m$.

@@

\begin{definition}
\label{def:noncrossingComplex}
We call \defn{non-crossing complex} of the pair of dual cellular dissections~$(\dissection, \dissection\dual)$ the simplicial complex~$\NCC = \AC = \SC$.
\end{definition}

\begin{remark}
To clearly visualize Proposition~\ref{prop:accordionsSlaloms}, we follow the convention to draw the $\dissection$-accordions (resp.~$\dissection\dual$-slaloms) so that they intersect the arcs of~$\dissection$ (resp.~of~$\dissection\dual$) only at the intersection points of the arcs of~$\dissection$ with the arcs of~$\dissection\dual$.
\vincent{That is interior points}
\end{remark}

\begin{remark}
The $\dissection$-slaloms and the $\dissection\dual$-accordions are defined dually and also coincide.
\end{remark}

\begin{remark}
A quick disclaimer about accordion complexes and slalom complexes of arbitrary dissections, non-necessarily cellular.
As stated in Proposition~\ref{prop:conditionsDualDissections}, a dissection~$\dissection$ with a face containing more than one boundary edge does not admit a dual cellular dissection~$\dissection\dual$.
This prevents to use the definitions given in this section.
Although its accordion complex (resp.~slalom complex) could still be defined as in~\cite{MannevillePilaud-accordion} (resp.~\cite{GarverMcConville}) for the disk, the resulting complexes would be join of smaller accordion complexes (resp.~slalom complexes) defined in this paper.
See \cite[Prop.~2.4]{MannevillePilaud-accordion} for a detailed statement in the case of the disk.
\end{remark}

@@

In this section, we show that non-kissing complexes and non-crossing complexes actually coincide.
Namely, any pair of dual cellular dissections~$(\dissection, \dissection\dual)$ defines a gentle bound quiver~$\bar Q_{\dissection, \dissection\dual}$ such that the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.
Conversely, any gentle bound quiver~$\bar Q$ gives rise to a surface~$\surface_{\bar Q}$ equipped with a pair of dual cellular dissections~$\dissection_{\bar Q}$ and~$\dissection\dual_{\bar Q}$ such that the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, \dissection\dual_{\bar Q}]$.
In fact, we show that the accordion, slalom, and non-kissing complexes can all be pictured in a unique way on the surface.
Interestingly, via the correspondence between pairs of dual cellular dissections and gentle bound quivers, the duality between the dissections~$\dissection$ and~$\dissection\dual$ translates to the Koszul duality between the gentle algebras of~$\bar Q$ and~$\bar Q\dual$.

@@

\subsection{Accordion, slalom and non-kissing complexe coincide}

@@

\begin{theorem}
\label{thm:complexesCoincide}
\begin{itemize}
\item For any gentle bound quiver~$\bar Q$, the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, \dissection\dual_{\bar Q}]$, therefore to the accordion complex~$\AC[\dissection_{\bar Q}]$ and to the slalom complex~$\SC[\dissection\dual_{\bar Q}]$.
\item For any pair of dual cellular dissections~$\dissection, \dissection\dual$ of an oriented surface, the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.
\end{itemize}
\end{theorem}
